Evapotranspiration mapping with Landsat 5TM in GRASS GIS

### 2011-01-04 tutorials/grass_landsat_ETa/article_GIPE.tex
fixed typos
git-svn-id: https://svn.osgeo.org/grass/grass-promo@44849 15284696-431f-4ddb-bdfa-cd5b030d7da7

%\setkeys{Gin}{width=1\textwidth}
\begin{figure}[htbp]
   \centering
   %name of your graphic, without the path AND in PNG (screnshots etc)/PDF (drawings) format:
   \includegraphics[scale=0.4]{gipe023.png}
   %caption of the figure
   \caption{r.sun Module}
   %label of the figure, which has to correspond to \ref{}:
   \label{fig:gipe023}
\end{figure}

\begin{center}
\begin{tabular}{lllll}
Landscapes & Mountains & Rural & City & Industrial\\
January & 1.5 & 2.1 & 3.1 & 4.1\\
February & 1.6 & 2.2 & 3.2 & 4.3\\
March & 1.8 & 2.5 & 3.5 & 4.7\\
April & 1.9 & 2.9 & 4 & 5.3\\
May & 2 & 3.2 & 4.2 & 5.5\\
June & 2.3 & 3.4 & 4.3 & 5.7\\
July & 2.3 & 3.5 & 4.4 & 5.8\\
August & 2.3 & 3.3 & 4.3 & 5.7\\
September & 2.1 & 2.9 & 4 & 5.3\\
October & 1.8 & 2.6 & 3.6 & 4.9\\
November & 1.6 & 2.3 & 3.3 & 4.5\\
December & 1.5 & 2.2 & 3.1 & 4.2\\
Annual & 1.9 & 2.75 & 3.75 & 5
\end{tabular}
\linebreak
Table 1: Linke Turbidity Coefficient
\end{center}

%\setkeys{Gin}{width=1\textwidth}
\begin{figure}[htbp]
   \centering
   %name of your graphic, without the path AND in PNG (screnshots etc)/PDF (drawings) format:
   \includegraphics[scale=0.4]{gipe024.png}
   %caption of the figure
   \caption{r.sun Module}
   %label of the figure, which has to correspond to \ref{}:
   \label{fig:gipe024}
\end{figure}

@@

\subsection{Sensible Heat Flux}
The Sensible heat flux (H) is calculated by "i.eb.h\_SEBAL01", given both maps of Net Radiation and soil Heat flux (Rn, G) at instantaneous time, the surface roughness (z0m), a map of the altitude corrected temperature (t0dem), a point data of the frictional velocity (u*), a value of actual vapour pressure (ea[KPa]) and the (x,y) pairs for wet and dry pixels.\newline
Surface roughness (Z0m) is an dependent variable of vegetation hight. Generation of an accurate map of the vegetation inside the study area in the date of image acquisition will lead to increase the accuracy of H calculation.\newline
\begin{smallverbatim}
    Z0m=hveg * 0.136 
\end{smallverbatim}
hveg is the hight of vegetation (m).\newline
Altitude corrected temperature map can be computed by using the toar thermal band and the DEM.\newline
\begin{smallverbatim}
    T0DEM=LST-(0.00627*DEM)
\end{smallverbatim}
The factor 0.00627 is the gradient of decreasing the temprature in relation to the decrease in hight (6.27 degree in each 1000 meter).\newline

The U* (Called U star) is the friction velocity is computed by:\newline
\begin{smallverbatim}
    U*= 0.41*um/(ln(hu/Z0m) 
\end{smallverbatim}
um is wind speed in the time of satellite overpass (m/s), hu is the hight of wind measurement which is normally described as 5 (m). Z0m here is not the same as that is used in  U* calculation. The  hveg in this case is defined as the hight of vegetation around the weather station.\newline

The value of actuale vapor pressure ea is calculated using the e0 for max and min tempreture.\newline
\begin{smallverbatim}
    e0=0.618*exp(17.27*T/(T+273.3))  
\end{smallverbatim}
e0 is the saturation vapor pressure at the air temperature. It must be calculated for Max and Min absolute tempreture.\newline
ea is calculated using:\newline
\begin{smallverbatim}
    ea=RH*(e0Max=e0Min)/2))/100  
\end{smallverbatim}
RH is the measured relative humidity at the time of satellite overpass.\newline
The location (row and column) of the wet (cold) and dry (hot) pixel must be found by the user. For wet pixel a place inside a water body and for dry pixel a place inside a bare and dry soil surface is recommended (SEBAL method is geographically dependent on extrem energy balance points).\newline

Sensible heat flux can be calculated with "i.eb.h\_SEBAL01".\newline

>>>>

%\setkeys{Gin}{width=1\textwidth}
\begin{figure}[htbp]
   \centering
   %name of your graphic, without the path AND in PNG (screnshots etc)/PDF (drawings) format:
   \includegraphics[scale=0.4]{gipe023.png}
   %caption of the figure
   \caption{r.sun Module, input options label}
   %label of the figure, which has to correspond to \ref{}:
   \label{fig:gipe023}
\end{figure}

Table 1: Linke Turbidity Coefficient
\begin{center}
\begin{tabular}{lllll}
Landscapes & Mountains & Rural & City & Industrial\\
January & 1.5 & 2.1 & 3.1 & 4.1\\
February & 1.6 & 2.2 & 3.2 & 4.3\\
March & 1.8 & 2.5 & 3.5 & 4.7\\
April & 1.9 & 2.9 & 4 & 5.3\\
May & 2 & 3.2 & 4.2 & 5.5\\
June & 2.3 & 3.4 & 4.3 & 5.7\\
July & 2.3 & 3.5 & 4.4 & 5.8\\
August & 2.3 & 3.3 & 4.3 & 5.7\\
September & 2.1 & 2.9 & 4 & 5.3\\
October & 1.8 & 2.6 & 3.6 & 4.9\\
November & 1.6 & 2.3 & 3.3 & 4.5\\
December & 1.5 & 2.2 & 3.1 & 4.2\\
Annual & 1.9 & 2.75 & 3.75 & 5
\end{tabular}
\linebreak
\end{center}

%\setkeys{Gin}{width=1\textwidth}
\begin{figure}[htbp]
   \centering
   %name of your graphic, without the path AND in PNG (screnshots etc)/PDF (drawings) format:
   \includegraphics[scale=0.4]{gipe024.png}
   %caption of the figure
   \caption{r.sun Module, optional label}
   %label of the figure, which has to correspond to \ref{}:
   \label{fig:gipe024}
\end{figure}

@@

\subsection{Sensible Heat Flux}
The Sensible heat flux (H) is calculated by "i.eb.h\_SEBAL01", given both maps of Net Radiation and soil Heat flux (Rn, G) at instantaneous time, the surface roughness (z0m), a map of the altitude corrected temperature (t0dem), a point data of the frictional velocity (u*), a value of actual vapour pressure (ea[KPa]) and the (x,y) pairs for wet and dry pixels.\newline
Surface roughness (Z0m) is an dependent variable of vegetation height. Generation of an accurate map of the vegetation inside the study area in the date of image acquisition will lead to increase the accuracy of H calculation.\newline
\begin{smallverbatim}
    Z0m=hveg * 0.136 
\end{smallverbatim}
hveg is the height of vegetation (m).\newline
Altitude corrected temperature map can be computed by using the toar thermal band and the DEM.\newline
\begin{smallverbatim}
    T0DEM=LST-(0.00627*DEM)
\end{smallverbatim}
The factor 0.00627 is the gradient of decreasing the temprature in relation to the increase in height (6.27 degree in each 1000 meter).\newline

The U* (Called U star) is the friction velocity is computed by:\newline
\begin{smallverbatim}
    U*= 0.41*um/(ln(hu/Z0m) 
\end{smallverbatim}
um is wind speed in the time of satellite overpass (m/s), hu is the height of wind measurement which is normally described as 5 (m). Z0m here is not the same as that is used in  U* calculation. The  hveg in this case is defined as the height of vegetation around the weather station.\newline

The value of actuale vapor pressure ea is calculated using the e0 for max and min tempreture.\newline
\begin{smallverbatim}
    e0=0.618*exp(17.27*T/(T+273.3))  
\end{smallverbatim}
e0 is the saturation vapor pressure at the air temperature. It must be calculated for Max and Min absolute tempreture.\newline
ea is calculated using:\newline
\begin{smallverbatim}
    ea=RH*(e0Max+e0Min)/2))/100  
\end{smallverbatim}
RH is the measured relative humidity at the time of satellite overpass.\newline
The location (row and column) of the wet (cold) and dry (hot) pixel must be found by the user. For wet pixel a place inside a water body and for dry pixel a place inside a bare and dry soil surface is recommended (SEBAL method is geographically dependent on extrem energy balance points).\newline
